### 2024-04-30 draft.tex
improve draft

\documentclass[a4paper,11pt]{amsart}

\usepackage{globaldef}
\input{definitions.tex}

\bibliography{bib-database.bib}

\usepackage{amsaddr} % affiliation in the first page

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\(Q\)-Boson model and relations with integrable hierarchies}

\author{Thiago Araujo}

\address{\noindent 
Instituto de Física Teórica, UNESP-Universidade Estadual Paulista, \\
Rua Dr. Bento T. Ferraz 271, Bl. II, São Paulo 01140-070, SP, Brazil
\\ \& \\
Instituto de Física, Universidade de São Paulo,\\
Rua do Matão, Travessa 1371, 05508-090 São Paulo, SP. Brazil
}

\email{\texttt{\href{tr.araujo@unesp.br}{tr.araujo@unesp.br}}}

\begin{abstract}
This work investigates the intricate relationship between the q-boson
model, a quantum integrable system, and classical integrable systems
such as the Toda and KP hierarchies. Initially, we analyze scalar
products of off-shell Bethe states and explore their connections to
tau functions of integrable hierarchies. Furthermore, we discuss
correlation functions within this formalism, examining their
representations in terms of tau functions, as well as their Schur
polynomial expansions.
\end{abstract}

\date{\today}
\keywords{Integrability, Bethe states, Schur, Hall-Littlewood, Toda, KP}
\subjclass[2020]{82B20, 82B23}

\maketitle

\setcounter{tocdepth}{1}
\tableofcontents

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Introduction}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

The exploration of connections between quantum and classical
exactly solvable models constitutes a significant endeavor aimed at
elucidating the overarching structure of integrable systems.  This
research avenue has yielded fruitful insights, as evidenced
by~\cite{Its:1992bj, Foda:2009zz, Alexandrov:2011aa, Araujo:2021ghu},
to name a few.  The present work is situated within this research field.

Here, we examine the emergence of classical integrable structures
within correlation functions of the q-boson system. This quantum
integrable system describes q-deformed bosons confined to
a one-dimensional chain~\cite{Bogoliubov:1992, Bogoliubov:1997soj,
  Bogoliubov2005}. Notably, this model bears close relation to the AL
(Ablowitz-Ladik) model, which is an integrable discretization of the
nonlinear Schrödinger equation.

In~\cite{Bogoliubov2005}, the author demonstrates that the \(q = 0\)
limit of the q-boson, known as the \emph{phase model}, is associated
with the enumeration of plane partitions. This connection allows for a
Schur polynomials expansion for the Bethe states. Subsequently, this
result was extended to \(q > 0\) in~\cite{Tsilevich:2006}, where the
Bethe states are expressed in terms of Hall-Littlewood polynomials.

The authors of~\cite{Foda:2008hn, Wheeler:2010vmq} utilize a rich
set of dualities between the ring of symmetric functions and
\(q\)-deformations of the Heisenberg algebra~\cite{Jing1991, Jing1995}
to explore the combinatorial properties of this spin chain. As a
consequence, it has been shown that the scalar product of off-shell
Bethe states are restricted tau functions of the KP
(Kadomtsev-Petviashvili) hierarchy.

This underlying combinatorial structure underscores the significance
of this model in various physical phenomena, notably in the context of
crystal melting and string theory~\cite{Okounkov:2003sp,
  Saulina:2004da}.  The application of the q-boson model in these
contexts has been demonstrated in~\cite{Sulkowski:2008mx}, with the
string data recoverable as the chain length approaches infinity.

This paper builds upon these developments and aims to extend some of
these results. Firstly, we investigate the connections between
the scalar product and tau functions, exploring them in a wider
context, particularly in the case of the Toda
hierarchy~\cite{Takasaki:2018wsv}. It is important to note that both
the AL and the KP hierarchies are reductions of the Toda
hierarchy. This paper takes steps towards resolving this problem,
which is instrumental in understanding the finer details of these
relationships.

Additionally, we explore other correlation functions within this model
and examine how they fit into the classical integrable context. It is
noteworthy that even when the object is not a tau function itself, it
may have an interesting expansion in terms of different types of
Schur polynomials, highlighting some of its combinatorial
properties\footnote{Results involving Schur and Hall-Littlewood
polynomials have been confirmed using the
software~\cite{Araujo:2024piv}.}. We also investigate some of these
expansions in this paper.

In Section 2, we provide a review of the phase model and
q-bosons. This section serves to establish notation and emphasize the
key aspects relevant to our analysis.  Section 3 delves into the
analysis of the phase model and its connections with the KP and Toda
integrable hierarchies. We explore the general expansion of a phase
model correlation function that yields a Toda system solution.  In
Section 4, we shift our focus to the q-bosons. We discuss a
determinantal formula for the scalar product of two off-shell Bethe
states and its expansion in terms of big, supersymmetric Schur
functions, as well as in terms of Kostka-Foulkes polynomials.
Finally, in Section 5, we conclude with an overview of open problems.

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Phase Model and q-Bosons}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

This section introduces the \emph{phase} and \emph{\(q\)-boson}
models~\cite{Bogoliubov:1992, Bogoliubov:1997soj, Bogoliubov2005,
  Tsilevich:2006}. It serves as a comprehensive review of existing
literature, offering insights into established concepts.  While the
content does not introduce novel ideas, its presentation adds value,
particularly from a pedagogical standpoint.  In structuring this
discussion, we adopt certain conventions from~\cite{Wheeler:2010vmq},
while aligning our presentation style more closely with that
of~\cite{Tsilevich:2006}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Phase model}
Consider the \((M+1)\) set of operators \(\{\phi_i,
\phi_i^\dagger,\mathcal{N}_i\}_{i=0}^M\) such that
\begin{equation}
 [\mathcal{N}_i, \phi_j] = - \phi_i \delta_{i,j} \quad
 [\mathcal{N}_i, \phi_j^\dagger] =  \phi_i^\dagger \delta_{i,j}  \quad 
 [\phi_i, \phi_j^\dagger] =  \pi_i \delta_{i,j}  
\end{equation}
where \(\pi_i =(\ket{0} \bra{0})_i\) is the vacuum projection operator.
These operators can be written as
\begin{equation}
\phi = \sum_{n\geq 0}\ket{n}\bra{n+1}\quad 
\phi^\dagger = \sum_{n\geq 0}\ket{n+1}\bra{n} \quad 
N = \sum_{n\geq 0}n\ket{n}\bra{n}\; ,
\end{equation}
and it is easy to see that \(\phi^\dagger \phi = \bm{1} - \ket{0}
\ket{0}\) and \(\phi\phi^\dagger = \bm{1}\).

The Hamiltonian is given by
\begin{equation}
  H = - \frac{1}{2} \sum_{n =0}^M \left(\phi_n^\dagger \phi_{n+1}
  + \phi_n \phi_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; .
\end{equation}
where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\) is the total
number operator, and we also impose periodic boundary conditions
\(\phi_{M+1} = \phi_0\) and \(\phi_{M+1}^\dagger = \phi_0^\dagger\).

These operators appear in the context of quantum optics, and for this
reason, this model is referred to as the \emph{phase model}. It corresponds to
the strongly correlated limit of the \(q\)-bosons
model~\cite{Bogoliubov:1997soj}, which we will define shortly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Representation}
The representation of the phase model algebra is constructed using
the vacuum state defined by \(\ket{0}_i\) by \(\phi_i\ket{0}_i =
0\). In this context, the state with \(n_i\) bosons (oscillators) is
given by \(\ket{n_i}_i = (\phi_i^\dagger)^{n_i} \ket{0}_i\).

Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
\otimes \cdots \otimes  \ket{0}_M\), the Fock space is defined as 
\begin{equation}
  \mathcal{F} = \bigotimes_{i=0}^M \mathcal{F}_i = \mathrm{span}
  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; ,
\end{equation}
where the states \(\ket{\vec{n}}\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 \equiv  (\phi_0^\dagger)^{n_0} (\phi_1^\dagger)^{n_1} \cdots  (\phi_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
with \(\phi_i^\dagger \equiv \bm{1} \otimes \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).  Moreover, it is easy
to see that these states are normalized, that is
\(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).  Finally,
the actions of the operators \(\mathcal{N}_i\) and \(\pi_i\) are
\begin{equation}
    \pi_i\ket{\vec{n}}  = \delta_{n_i, 0} \ket{\vec{n}} \qquad 
    \mathcal{N}_i\ket{\vec{n}} = n_i \ket{\vec{n}}\; .
\end{equation}

Given a state \(\ket{\vec{n}} = \ket{n_0, n_1, \dots, n_M}\), we
associate a partition \( \lambda = (1^{n_1} 2^{n_2} \cdots
M^{n_M})\).  It's worth noting that this correspondence is not unique,
as the partition \(\lambda\) does not account for the number of
particles \(n_0\).  If the total number of particles \(N\) is known,
we can determine \(n_0 = N - \ell(\lambda)\), where \(\ell(\lambda)\)
represents the number of rows in the Young diagram defined by
partition \(\lambda\).  This aspect is crucial because, in our
subsequent considerations, the \(N\) particle sector remains fixed
owing to the integrability of the model. Consequently, the value of
\(n_0\) becomes known once we specify the partition \(\lambda\).

Finally, based on the correspondence we mentioned in the introduction,
Wheeler~\cite{Wheeler:2010vmq} defines a map \(\mathcal{M}_\psi:
\mathcal{F} \to \mathcal{F}^{(0)}_\psi\), where
\(\mathcal{F}^{(0)}_\psi\) denotes the Fock space of charged free
fermions constructed from the neutral (fermionic) vacuum.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Bethe Ansatz}
The \(L\)-matrix is given by 
\begin{equation}
  L_{an} = 
\begin{pmatrix}
x^{ - 1/2} & \phi_n^\dagger \\ \phi_n & x^{1/2}
\end{pmatrix}_a\; , 
\end{equation}
where \(x \in \mathbb{S}^1\). We also have the monodromy matrix
\begin{equation}
  T_a(x) = L_{aM}(x) \cdots L_{a0}(x) = 
\begin{pmatrix}
A(x) & B(x) \\ C(x) & D(x)
\end{pmatrix}_a\; .
\end{equation}

With these expressions, one can finally build the Bethe states
\begin{equation}
  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \qquad 
  \bra{\Psi(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j) 
\end{equation}
where \(\mathbb{B}(y) = y^{M/2} B(y)\) and \(\mathbb{C}(y) = y^{M/2} C(1/y)\).

When the coordinates \(\{ y_j \ | \ j =1, \dots , N\}\) satisfy the Bethe equations
\begin{equation}
\label{eq:bethe_eq}
  y^{N + M}_i = (-1)^{N-1} \prod_{\substack{j = 1 \\ j \neq i}}^N y_j\; , \qquad i = 1, \dots, N\; , 
\end{equation}
we say that the Bethe states are \emph{on-shell}; otherwise, we have
\emph{off-shell} states. In what follows, we will only consider
\emph{off-shell} states.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{q-Bosons}
The set of operators \(\{b_i, b_i^\dagger,\mathcal{N}_i\}_{i=0}^M\) constitute
\(M+1\) independent q-boson algebras defined by
\begin{equation}
[\mathcal{N}_i, b_j^\dagger]=\delta_{i,j} b_i^\dagger\; , \quad 
[\mathcal{N}_i, b_j]=-\delta_{i,j}b_i\; , \quad
[b_i, b_j^\dagger]= \delta_{i,j} q^{-2\mathcal{N}_i}  \equiv \delta_{i,j} Q^{\mathcal{N}_i}\; , 
\end{equation}
where we denote the deformation parameter as \(Q = q^{-2}\).

The q-boson model is characterized by its Hamiltonian
\begin{equation}
  \mathcal{H} = -\frac{1}{2} \sum_{n=0}^M
  \left(b_n^\dagger b_{n+1} + b_n b_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
\end{equation}
where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and we also
impose periodic boundary conditions \(b_{M+1} = b_0\) and
\(b_{M+1}^\dagger = b_0^\dagger\).

In the limit \(Q \to 1\), the q-bosons behave as ordinary bosons,
while the limit \(Q \to 0\) (\(q \to \infty\)) corresponds to the
phase model discussed earlier.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Representation}
Let us define the \(i\)-th vacuum \(\ket{0}_i\) such that
\(b_i\ket{0}_i = 0\). The representation of this Hilbert space,
denoted by \(\mathcal{F}_i^Q\), is defined by the states
\(\ket{n_i}_i \propto (b_i^\dagger)^{n_i} \ket{0}_i\).

Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
\otimes \cdots \otimes  \ket{0}_M\),
the Fock space is defined as 
\begin{equation}
  \mathcal{F}_{Q} = \bigotimes_{i=0}^M \mathcal{F}_i^Q = 
  \mathrm{span}\left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; ,
\end{equation}
where the actions of the operators \(\{b_i, b_i^\dagger\}\) are given
by the following relations
\begin{subequations}
\begin{alignat}{3}
    & b_0\ket{n_0} =(1 - \delta_{0, n_0}) \frac{(1 - Q^{n_0})}{(1 - Q)^{1/2}}\ket{n_0 - 1}
    &\hspace{0.5cm}& b_0^\dagger \ket{n_0} =  \frac{1}{(1 - Q)^{1/2}}\ket{n_0 + 1}  \\
    & b_i\ket{n_i} = \frac{(1 - \delta_{0, n_i})}{(1 - Q)^{1/2}}\ket{n_i - 1}
    && b_i^\dagger \ket{n_0} =  \frac{(1 - Q^{n_i+1})}{(1 - Q)^{1/2}}\ket{n_0 + 1}\quad i\neq 0\; .
\end{alignat}
\end{subequations}

Consequently, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (b_0^\dagger)^{n_0} (b_1^\dagger)^{n_1} \cdots  (b_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(b_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
b_i^\dagger \otimes \cdots \otimes \bm{1}\).
Moreover, this space has an inner product that satisfies
\begin{equation}
  \bracket{\vec{n}}{\vec{m}} = \frac{[m_0]!}{\prod_{i=1}^M[m_i]!} \delta_{\vec{n}\vec{m}}\qquad 
    [n]! =
    \left\{
    \begin{array}{ll}
    \prod_{i=1}^n(1 - Q^i) & \textrm{if}\ \ n\neq 0 \\
    1 & \textrm{otherwise}
  \end{array}\right.\; .
\end{equation}

Similar to the phase model, we can associate to a given states
\(\ket{\vec{n}} = \ket{n_0, n_1, \dots, n_M}\), a Young diagram
\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\). It is useful to define
a proportionaly factor relating these two objects
\begin{equation}
  \ket{\lambda}_Q  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
\end{equation}
where \(p_i(\lambda)\) denotes the number of parts of size \(i\) in
the partition. These partition states satisfy
\(\bracket{\lambda}{\mu}_Q = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
again, we see that the correspondence is not unique, as the number of
oscillators at site \(i=0\) is completely ignored in the partition
notation. Finally, if the number of particles is fixed, then \(n_0 = N
- \ell(\lambda)\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Bethe Ansatz}
The \(L\)-operator for the q-boson is given by
\begin{equation}
  L_{an}(x, Q) =
  \begin{pmatrix}
    x^{-1/2} & (1 - Q)^{\frac{1}{2}} b_n^\dagger \\ (1 - Q)^{\frac{1}{2}} b_n & x^{1/2}
  \end{pmatrix}_a\; ,
\end{equation}
and the monodromy matrix is 
\begin{equation}
  T_a(x,Q) = L_{am}(x, Q)  \dots  L_{a0}(x, Q) = 
  \begin{pmatrix}
    A(x, Q) & B(x, Q) \\ C(x, Q) & D(x, Q)
  \end{pmatrix}_a\; .
\end{equation}

As before, the eigenstates of the Hamiltonian have the form
\begin{equation}
  \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N \mathbb{B}(y_j, Q) \ket{\bm{0}}\qquad 
  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j, Q) \; .
\end{equation}
where \(\mathbb{B}(y, Q) = y^{M/2} B(y, Q)\) and \(\mathbb{C}(y, Q) =
y^{M/2} C(1/y, Q)\). When the parameters \(\{ y_j \ | \ j =1, \dots , N\}\)
satisfy the Bethe equations given by
\begin{equation}
  y^{N + M}_i =\prod_{\substack{j = 1 \\ j \neq i}}^N\frac{Q y_i - y_j}{y_i - Q y_j}\; , \qquad i = 1, \dots, N\; , 
\end{equation}
we have \emph{on-shell} Bethe states; otherwise, they are \emph{off-shell} states.

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Tau functions in the phase model}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

This section explores the presence of integrable hierarchies in the
phase model. We demonstrate its relations with the Toda hierarchy tau
function and discuss some implications, particularly its connection to
a matrix model. We also argue that these results imply that the scalar
products in the model also serve as KP hierarchy tau functions,
consistent with the findings of
Wheeler~\cite{Wheeler:2010vmq}. Additionally, we highlight the
existence of correlation functions in this model that satisfy the KP
hierarchy equations.

Bogoliubov~\cite{Bogoliubov2005} has demonstrated that the scalar
product of two vectors in the \(N\)-particle sector of a chain with
length \(M+1\) is
\begin{equation}
\begin{split}
\label{eq:scalar}
  \mathcal{I}(N, M|\bm{x}, \bm{y}) & =
  \bra{\bm{0}} \prod_{i=1}^N \mathbb{C}(x_i) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
 & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
\end{split}
\end{equation}
where \(H\equiv [H_{ij}]_{i,j=1}^N\) is an \(N\times N\) matrix with components
\begin{equation}
\label{eq:h-matrix}
  H_{ij} = H(x_i, y_j) 
  =\frac{1 - (x_i y_j)^{ M + N}}{1 - x_i y_j }\; .
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Toda tau functions}
We now argue that the scalar product defined above is a tau function
of the Toda hierarchy. Let us first write the function \(H(z,w)\) as
the geometric sum
\begin{equation}
\label{eq:h-exp}
  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; .
\end{equation}
Hence, we express the determinant of the \(H\) matrix as
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) = \det_{i,j} \left( \sum_{k=1}^{M+N} x_i^{k-1} y_j^{k-1}\right)
\end{equation}
Furthermore, we can interpret this expression as the result of
multiplying an \(N\times (N+M)\) matrix \(\mathcal{X}\) by another
\((M + N)\times N\) matrix \(\mathcal{Y}\), which are given by
\begin{equation}
  \mathcal{X} = 
  \begin{pmatrix}
  x_1^0 & x_1^1 & \dots & x_1^{M+N-1} \\  
  x_2^0 & x_2^1 & \dots & x_2^{M+N-1} \\  
  \vdots \\
  x_N^0 & x_N^1 & \dots & x_N^{M+N-1} 
  \end{pmatrix}\quad \textrm{and} \quad 
  \mathcal{Y} = 
  \begin{pmatrix}
  y_1^0 & y_2^0 & \dots & y_N^0 \\  
  y_1^1 & y_2^1 & \dots & y_N^1 \\  
  \vdots & \vdots & & \vdots \\
  y_1^{N+M-1} & y_2^{M+N-1} & \dots & y_N^{M+N-1}
  \end{pmatrix}\; ,
\end{equation}
therefore 
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right)
  \equiv  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
  = \sum_{0 \leq \ell_{N} \leq \dots \leq \ell_1\leq N+M } \det_{ik}(x_i^{\ell_k}) \det_{ik}(y_j^{\ell_k})\; .
\end{equation}
If we now define \(\ell_j = \lambda_k - k + N\), and
using~(\ref{eq:scalar}), we have
\begin{equation}
\label{eq:scalar_exp}
\mathcal{I}(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
\frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; .
\end{equation}
where \(\Delta(\bm{x})\) and \(\Delta(\bm{x})\) are Vandermonde
determinants. This formula agrees with the Schur expansion defined
in~\cite{Bogoliubov2005}.

Define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\) and
\(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
\begin{equation}
  t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; .
\end{equation}
One can write the inner product in terms of these coordinates, that is
\begin{equation}
  \mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
This expression is known to be a tau function for \(M, N\to \infty\).
As such, utilizing the free fermions representation~\cite{Alexandrov:2012tr}, it can be written as
\begin{equation}
  \lim_{M,N\to \infty}\mathcal{I}(N, M|\bm{t}, \bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}
It is a tau function of the Toda hierarchy with trivial element
\(\mathbb{1} \in GL(\infty)\), and it is nothing but the Cauchy's
identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_{-m} \right) \; .
\end{equation}

Bringing all these facts together, the truncation for finite \(M\) and
\(N\) also yields tau functions of the Toda hierarchy. More
specifically, according to~\cite{Alexandrov:2012tr, Kharchev:1991gd,
  Zabrodin:2010ii}, the truncation of the tau function corresponds to
the inclusion of a projection operator in the expectation value of the
tau function written in the fermionic representation.

As a final remark, we can also write
\begin{equation}
  H(z,w) = \sum_{k=0}^{M+N-1} h_k(zw)^{k} \; , 
\end{equation}
with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
case, one can define a diagonal \((N+M)\times (N+M)\) matrix
\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Consequently,
if we repeat the arguments above, we find
\begin{equation}
\mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda \subseteq [N,M]} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
Here, \(h_\lambda\) is equal to \(1\) if \(\lambda \in [N,M]\), and it
is zero otherwise.

In this case, we find that this tau function is a trivial example of
the tau functions considered in~\cite{orlov:2001}. We anticipate that
in the analysis of more general correlation functions, the diagonal
terms \(h_\lambda\) will be more interesting. We will revisit this
discussion soon.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Matrix Models}
It is also interesting to note the particular case when \(M\to
\infty\), but with a finite number of particles \(N\). In this case,
the partitions \(\lambda \in [N, \infty]\) satisfy the condition
\(\ell(\lambda) \leq N\). Therefore, the scalar
product~(\ref{eq:scalar_exp}) becomes
\begin{equation}
  \mathcal{I}_N(\bm{t}, \bm{t}')\equiv 
  \lim_{M\to \infty}\mathcal{I}(N, M|\bm{t}, \bm{t}')
  = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}} s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
\end{equation}
From~\cite{Zabrodin:2010ii}, we know that this expression can be
written as the following integral
\begin{equation}
  \mathcal{I}_N(\bm{t}, \bm{t}') =
  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
  e^{\xi(\bm{t}, z_\ell) - \xi(\bm{t}', z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})\; ,
\end{equation}
where
\begin{equation}
\xi(\bm{t}, z) = \sum_{k\geq 0} t_k z^k \qquad
\xi(\bm{t}', 1/z) = \sum_{k\geq 0} t'_{-k} z^{-k} \; . 
\end{equation}
See also~\cite{Kharchev:1991gd} for a detailed proof of this relation,
and~\cite{Orlov:2005} for other details.

From the results of~\cite{Zabrodin:2010ii}, see citations therein, we
have an interesting consequence of this representation. Impose the
Bethe equations~(\ref{eq:bethe_eq}) to one set of variables, say \(x_j
= e^{-ip_j} \in \mathbb{S}^1\). Additionally, let us set \(\bm{t}' = -
\bm{t}^\star\). In this particular case, we have
\begin{equation}
\xi(\bm{t}, z) - \xi(\bm{t}', 1/z)  = 2 \ \textrm{Re}\left(\sum_k t_k z^k\right)\; .
\end{equation}
Then, the phase model is equivalent to an ensemble of \(N\) 2D Coulomb
particles on a circle. In this case, we find that the quantities
\(z_{\ell}\) are eigenvalues of a matrix \(U\).

Furthermore, according to Zabrodin~\cite{Zabrodin:2010ii}, see
citations therein, we also know that under the rescaling \(t_k \to
T_k/ \hbar\), \(t'_k \to T_{-k}/ \hbar\) and \(N = T_0/ \hbar\), we
obtain the dispersion limit tau function
\begin{equation}
  F_0(\bm{T}) = \log \mathcal{I}_{T_0}(\bm{T}, \bm{T}') + \mathcal{O}(\hbar)\; ,
\end{equation}
that is a free energy, from the viewpoit of the matrix integral
partition function.

It remains unclear how one can use this fact to determine properties
of the integrable model, but it might be possible to study the
analytic structure of the free energy \(F_0\) to gain some
understanding of the Bethe roots \(\bm{x}\). This problem is currently
under further investigation, and we hope to report new results
elsewhere.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{KP tau function}
Lastly, one may also observe that if we fix one set of coordinates,
say \(\bm{y}\), then we can write
\begin{equation}
 \mathcal{I}(N,M|\bm{t}) 
 = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i
  +j}(\bm{y}))\). But now, it is trivial to notice that these are
Plücker coordinates in the Jacobi-Trudi form, as seen
in~\cite{Miwa2000, Alexandrov:2012tr}.

Hence, the following expression
\begin{equation}
 \mathcal{I}(N,M|\bm{t}) = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
is also a KP tau function, a fact that we already know
from~\cite{Wheeler:2010vmq}, where the author proved this statement
using the free fermions formalism.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Correlation functions}
Bogoliubov has also shown, in~\cite{Bogoliubov2005}, that the
correlation functions
\begin{subequations}
\begin{equation}
\begin{split}
\label{eq:correlation}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \bra{0} \prod_{j=1}^N \mathbb{C}(x_j)
  \prod_{k=1}^{N-1} \mathbb{B}(y_k) \phi_m^\dagger \ket{0}\\
  & =  \prod_{j=1}^N x_{j}^{M/2} \prod_{k=1}^{N-1} y_{j}^{M/2}
  \bra{0} \prod_{j=1}^N C(1/x_j) \prod_{k=1}^{N-1} B(y_k) \phi_m^\dagger \ket{0}
\end{split}
\end{equation}
can be written as
\begin{equation}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\}) = 
  \frac{(-1)^{N-1}}{y_N^{(N-1)/2}} \prod_{j=1}^N x_{j}^{M/2}
  \prod_{k=1}^{N-1} y_{j}^{M/2}
  \left( \prod_{t<N} \frac{y_N - y_t}{y_t} \right)
  \frac{\det Q}{\det H} \mathcal{I}(N,M|\bm{x}, \bm{y})
\end{equation}
\end{subequations}
where \(Q\) is an \(N\times N\) matrix with components 
\begin{equation}
 Q_{jN} = x_j^{(M + N - 1- 2m)/2} \quad  \textrm{and} \quad 
 Q_{jk} = H_{jk} \; , 
\end{equation}
and \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\)
in~(\ref{eq:h-matrix}).  The components \(Q_{jN}\) are independent of
the coordinates \(y\); therefore, we cannot express the above
expression as a Toda hierarchy tau function.

We already know from~(\ref{eq:scalar}) that 
\begin{equation}
    \frac{\mathcal{I}(N,M|\bm{x}, \bm{y})}{\det H}=  \frac{1}{\Delta(x)\Delta(y)} \; ,
\end{equation}
then
\begin{equation}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})  \propto
  \frac{\det Q}{\Delta(x)} \equiv \mathcal{A}_m(N,M|\bm{x}, \bm{y})\; ,
\end{equation}
and we treat the coordinates \(\{ \bm{y} \}\) as a set of \(N\) fixed
parameters.

Furthermore, we define vector field \(\bm{F}(z) = (F_1, \dots, F_N)\),
where its components are given by
\begin{equation}
    F_j (z) = H(z, y_j) \quad \textrm{if} \ j \neq N\; , \qquad \textrm{and}\qquad 
    F_N (z)  = z^{(M + N - 1 - 2m)/2} \; .
\end{equation}
As before, we expand \(F_j(z)\), \(j\neq N\), as the geometric sum
\begin{equation}
  F_j(z) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n
  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n \; , \quad  f_{j, n} = y_j^n\; .
\end{equation}
We can also express the function \(F_N(z)\) in this form by setting
\(f_{N,n} = \delta_{n,(M - N - 1 - 2m)/2}\) and requiring \((M - N -
1)\) to be an even integer.

With these definitions, we conclude that
\begin{equation}
\mathcal{A}_m(N,M|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; .
\end{equation}
From the expansion 
\begin{equation}
  \det_{jk} F_j(x_k) = \det_{jk} \left(  \sum_{n=0}^{M + N -1} f_{j, n} x_k^n \right) \; ,
\end{equation}
and utilizing the Cauchy-Binet formula, we get
\begin{equation}
    \mathcal{A}_m(N,M|\bm{x}, \bm{y})
  = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
  \frac{\det_{jk}(f_{j, \ell_k}) \det_{jk}(x_k^{\ell_j})}{\Delta(x)}\; . 
\end{equation}

We now use the definition of the Schur polynomials and the
Jacobi-Trudi expression of Plücker coordinates, as detailed
in~\cite{Alexandrov:2012tr}, leading to the following expression:
\begin{equation}
\mathcal{A}_m(N,M|\bm{x}, \bm{y}) =
\sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
\end{equation}
where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\), \(\ell_k =
\lambda_k - k +N\).  Putting all these facts together, we conclude
that this expression is also a KP tau function.

This expression underscores the non-trivial nature of these tau
functions within the model. However, it also implies the existence of
other intriguing examples awaiting exploration. Let us now briefly
investigate other cases.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Skew Schur polynomials expansion}
From the coordinate expansion of the Bethe
vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
\begin{equation}
\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
 \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
\end{equation}
where \(s_{\mu/\lambda}\) are skew Schur polynomials~\cite{Macdonald:1998}.
Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
\begin{equation}
\begin{split}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
\end{split}
\end{equation}
where we have used that \(\phi^\dagger \ket{\bm{0}} = \ket{(m)}\), and in the second line
we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
This expression also shows that the expression \(A_m\) can also be written as a tau function.

More generally, let us consider the correlation functions 
\begin{equation}
\begin{split}
  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
  \bra{\lambda_1} \prod_{j=1}^N \mathbb{C}(x_j)
  \prod_{k=1}^{N'} \mathbb{B}(y_k)\ket{\lambda_2} \\
  & = \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; .
\end{split}
\end{equation}
where \(N' + \ell(\lambda_2) + n_0 = N' + \ell(\lambda_2) + n'_0\) and
we have also used that the skew Schur polynomial for any Young diagram
\(\mu\) that does not contain \(\lambda_1\) and \(\lambda_2\)
vanishes. Using the elementary property of skew Schur
polynomials~\cite{Macdonald:1998}
\begin{equation}
  \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) = \prod_{i,j}\frac{1}{1 - x_i y_j}
 \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; ,
\end{equation}
and the Cauchy's identity, we have 
\begin{equation}
\begin{split}
  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
 \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
 \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; .
\end{split}
\end{equation}
Therefore, this correlation function is nothing but the off-shell
norm~(\ref{eq:scalar}) times a finite sum over skew Schur
functions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Schur polynomials expansion of some correlation functions}
Indeed, it is worth noting that other quantities, which are not tau
functions, may have interesting Schur polynomial expansions. Consider
the state calculated in~\cite{Bogoliubov2005}.
\begin{equation}
  \ket{\mathcal{Y}} =
 \sum_{\vec{n}} \prod_{j=0}^M \ket{n_j} = \sum_\mu \ket{\mu} \qquad \sum_j n_j = N\; .
\end{equation}
Then
\begin{equation}
 \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\nu} = 
  \sum_{\lambda\subseteq [N,M]} s_{\lambda/\nu}(\bm{x})\; . 
\end{equation}
While it may not be a tau function, it has an interesting expansion as
a sum of skew Schur polynomials.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{General tau functions}
Based on the discussion we have had so far, and on the general mapping
between the Phase model and free fermions, one can grasp the general
form of tau functions in the context of this phase model.

Let us consider the vertex operator construction~\cite{Okounkov2001},
as also discussed in~\cite{Alexandrov:2012tr, Wheeler:2010vmq}. We
consider the vacuum state \(\ket{\bm{0}}\), often referred to as a
``Dirac sea'', defined by the conditions \(\psi_m\ket{\bm{0}} =
\psi_n^\star\ket{\bm{0}} =0 \) for \(m<0\) and \(n \geq 0\), where
\(\psi_n\) are components of a holomorphic free fermionic field. In
this formalism, the partition states are given by
\begin{equation}
  \ket{\mu} = \textrm{sign}(\sigma) \prod_{j=1}^d \psi_{a_j} \psi^\star_{-b_j}\ket{\bm{0}}\; ,
\end{equation}
where \(\textrm{sign}(\sigma) = \pm 1\) are defined in a such a way
that the Shur coefficients of the vertex operators (defined below)
have positive coefficients.

The pairs \(\{(a_j|b_j)\}_{j=1}^d\) define the Frobenius notation of
the partition \(\mu = (\mu_1, \mu_2, \dots, \mu_\ell)\). In this
notation, \(a_j\) is given by \(\mu_j - j\) and \(b_j\) is given by
\(\mu'_j - j\), where \(d\) represents the number of boxes in the
diagonal of the Young diagram, and \(\mu'\) is its conjugate, or
transpose, diagram. From these definitions, we have the equivalence
\begin{equation}
  \prod_{k\geq 1} (\phi_k^\dagger)^{n_k} \mapsto \textrm{sign}(\sigma) \prod_{j=1}^d
  \psi_{a_j} \psi^\star_{-b_j} \qquad |\mu| = \sum_k k n_k = \sum_j(a_j + b_j) + d\; .
\end{equation}

Finally, the \(\mathfrak{gl}(\infty)\) algebra has generators given by the
bilinears \(X = \sum_{j, j \in \mathbb{Z}} x_{ij} \bm{\colon} \psi_i \psi_j^\star\bm{\colon}
+ c\), where \(c\in \mathbb{C}\), \(x_{ij} =
0\) for large \(|j -i|\), say \(\geq M\), and the colons denote the normal ordering
\begin{equation}
  \bm{\colon} \psi_i \psi_j^\ast \bm{\colon} =  \psi_i \psi_j^\ast
  - \bra{0} \psi_i \psi_j^\ast \ket{0} \; .
\end{equation}
The group \(GL(\infty)\) is defined through the \(\exp:
\mathfrak{gl}(\infty) \to GL(\infty)\) as usual.

Note that these elements have finitely many non-zero entries,
including diagonal terms, upper triangular terms, and lower triangular
terms. These are equivalent to the number operators \(\mathcal{N}\),
annihilation operators \(\phi\), and creation operators
\(\phi^\dagger\), respectively. The central charge is equivalent to
the vacuum projection \(\pi = \ket{\bm{0}}\bra{\bm{0}}\). Therefore,
\begin{equation}
   GL(\infty) \ni G \mapsto
   \mathcal{G} = \exp \left[\sum_{i=1}^M \left( \sum_{a=1}^3
   c_{i, a} T_i^{a}  + c \pi_i \right)\right]\; ,
\end{equation}
where \(T^{a}_i \in \{ \mathcal{N}_i, \phi_i, \phi_i^\dagger \}_{i=1}^M\).

The operators \(\mathbb{B}(x)\) and \(\mathbb{C}(x)\), 
for a large enough chain \(M\to \infty\),
are related to the vertex operators
\(\Gamma_-(x)\) and \(\Gamma_+(x)\), respectivelly, as
\begin{equation}
  \begin{split}
    \mathbb{B}(x) \mapsto \Gamma_-(x) & = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{-n}\right) \\
    \mathbb{C}(x) \mapsto \Gamma_+(x) & = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{n}\right) \; ,
  \end{split}
\end{equation}
where \(J_n \) is written in terms of free fermions as \(J_n =
\sum_{j\in \mathbb{Z}} \bm{\colon} \psi_j \psi_{j+n}^\ast
\bm{\colon}\). These set of operators generate a Heisenberg subalgebra
\(\widehat{\mathfrak{gl}}(1) \subset \mathfrak{gl}(\infty)\)
\begin{equation}
  [J_m, J_n] = m \delta_{n+m,0}\; .
\end{equation}

Putting all these facts together, we have that the tau functions of the Toda 
hierarchies, given by
\begin{equation}
  \tau_s(\bm{x}, \bm{y}) = \bra{s} \prod_{i} \Gamma_+(x_i) G \prod_j \Gamma_-(x_j) \ket{s}\; ,
\end{equation}
are mapped into objects of the form 
\begin{equation}
\label{eq:tau-corr}
\begin{split}
  \tau(\bm{x}, \bm{y}) & = \bra{\Psi(\bm{x})} \mathcal{G} \ket{\Psi(\bm{y})} \\
  & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
  \mathcal{G} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\; ,
\end{split}
\end{equation}
where we necessarily have \(s=0\) in the phase model

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Hypergeometric tau functions}
From these expressions, we can conclude that if we consider a diagonal
group element \(\mathcal{G} = \exp \left( \sum_{i\geq 0} c_i
\mathcal{N}_i\right)\), we have that~(\ref{eq:tau-corr}) becomes
\begin{equation}
\begin{split}
  \tau(\bm{x}, \bm{y}) 
  & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
  e^{\sum_i c_i \mathcal{N}_i} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\\
  & = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu\nu} s_\mu(\bm{x}) s_\nu(\bm{x})
\end{split}
\end{equation}
where
\begin{equation}
  c_{\mu \nu} = \bra{\mu} e^{\sum_i c_i \mathcal{N}_i}\ket{\nu} \; .
\end{equation}
From the representation of the phase model, we have that for the partition
\(\mu = (1^{n_1} 2^{n_2} \dots M^{n_M})\), we have
\begin{equation}
  \mathcal{N}_i \ket{\mu} = n_i \ket{\mu} \qquad n_0 = N - \ell(\mu)\; .
\end{equation}

Consequently, we have
\begin{equation}
  c_{\mu \nu} \equiv \delta_{\mu\nu} c_{\mu}  =  \delta_{\mu\nu} \prod_i e^{c_i n_i}\; , 
\end{equation}
and we conclude that 
\begin{equation}
  \tau(\bm{x}, \bm{y}) 
  = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu} s_\mu(\bm{x}) s_\mu(\bm{x})\; ,
\end{equation}
We observe that this tau function has diagonal coordinates
\(c_\lambda\). These tau functions belong to the hypergeometric type
considered in~\cite{Orlov:2000, orlov:2001, Orlov:2001b, Orlov:2005}.

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Tau functions in the Q-bosons model}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

We now shift our focus to the case of q-bosons. The analysis parallels
what we did before, but the specific details are markedly
different. We begin by examining the norm of two off-shell Bethe
states.

It has been shown~\cite{Tsilevich:2006} (see
also~\cite{Sulkowski:2008mx, Wheeler:2010vmq}) that the Bethe states
\(\ket{\Psi(\bm{x})}\) have coordinate expansions
\begin{equation}
  \ket{\Psi(\bm{x})} = \sum_{\mu \subseteq [N,M]} P_\mu(\bm{x}, Q) \ket{\mu}_Q  \qquad 
  \bra{\Psi(\bm{x})} = \sum_{\mu \subseteq [N,M]} P_\mu(\bm{x}, Q) \prescript{}{Q}{\bra{\mu}}\; ,
\end{equation}
where \(P_\lambda\) denote Hall-Littlewood polynomials. In this form,
the scalar product of two off-shell Bethe states in the q-boson model
can be easily calculated to be
\begin{equation}
\label{eq:scalar-hl}
\mathcal{I}_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j, Q)
\prod_{k=1}^N \mathbb{B}(y_k, Q) \ket{\bm{0}}
= \sum_{\lambda \subseteq [N,M]} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
\end{equation}
where we use that \(\bracket{\lambda}{\mu} = b_\lambda(Q)
\delta_{\lambda, \mu}\) and the completeness relation
\begin{equation}
  \sum_{\lambda} \frac{1}{b_\lambda(Q)} \ket{\lambda}_Q {}_Q\bra{\lambda}  = \mathbb{1}\; .
\end{equation}

It turns out that this expansion is the Cauchy identity for
Hall-Littlewood polynomials~\cite{Macdonald:1998}
\begin{equation}
\label{eq:cauchy_hl}
\sum_{\lambda} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
= \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k}\; ,
\end{equation}

Our goal is to explore some properties of Hall-Littlewood polynomials
to gain insight into aspects of this expansion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Scalar product: determinant formula}
We aim to refine the expressions above. First, we consider a
determinant expression for the scalar product~(\ref{eq:scalar-hl}). We
proceed with the scenario where \(M\) and \(N\) are very large but
finite. In this case, we express this scalar product as
\begin{equation}
\label{eq:q-inner}
  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
= \prod_{j_1, j_2} (1-Q x_{j_1} y_{j_2}) \prod_{k_2, k_2}(1 - x_{k_1} y_{k_2})^{-1}\; .
\end{equation}

Using the Cauchy's identity for Schur polynomials, that is 
\begin{equation}
  \sum_\lambda s_\lambda(\bm{x}) s_\lambda(\bm{y}) = \prod_{i,j} \frac{1}{1 - x_i y_j}\; ,
\end{equation}
and from the results derived in the phase model, we find that
\begin{equation}
  \prod_{i,j}\frac{1}{1 - x_i y_j}  \equiv \mathcal{I}(N,M|\bm{x}, \bm{y}) = 
  \frac{\det_{j,k}H(\bm{x},\bm{y})}{\Delta(\bm{x}) \Delta(\bm{y})}\; .
\end{equation}
where \(H\) is the matrix~(\ref{eq:h-matrix}). Consequently, the scalar
product~(\ref{eq:q-inner}) becomes
\begin{equation}
  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
= \frac{\mathcal{I}(N,M|\bm{x}, \bm{y})}{\mathcal{I}(N,M|\bm{x}, Q\bm{y})}\; , 
\end{equation}
and we see that it is the quotient of scalar products of the phase
model. Hence, it is the quotient of two Toda tau functions.
Observe that in the case \(Q = 0\), we have \(\mathcal{I}(N, M|
\bm{x}, \bm{0}) = 1\), then \(\mathcal{I}_0(N,M | \bm{x}, \bm{y})
= \mathcal{I}(N,M | \bm{x}, \bm{y})\), as expected.

Additionally, we write
\begin{equation}
  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
  = \frac{\Delta(Q\bm{y})}{\Delta(\bm{y})}
  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x},Q \bm{y})} \; .
\end{equation}
Using \(\Delta(Q\bm{y})/ \Delta(\bm{y}) = \prod_{j=1}^{N-1} Q^j\), we have 
\begin{equation}
  \mathcal{I}_Q(N,M | \bm{x}, \bm{y}) =  Q^{N(N-1)/2}\;
  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x}, Q \bm{y})} \; .
\end{equation}
Let us denote \(H(\bm{x}, Q\bm{y}) = H_{Q}\), therefore
\begin{equation}
  \mathcal{I}_Q(N,M | \bm{x}, \bm{y}) = Q^{N(N-1)/2} \det H_{Q}^{-1} \det H
= Q^{N(N-1)/2} \det \mathcal{H}(\bm{x}, \bm{y}) \; ,
\end{equation}
where \(\mathcal{H} = H_{Q}^{-1} H\) is a matrix with components 
\(\mathcal{H}_{i,j} \equiv \mathcal{H}(x_i, y_j)\) where
\begin{equation}
  \mathcal{H}(z, w) = \sum_{k} \frac{(1 - z y_k)}{(1 - Q z y_k)} 
\frac{(1 - Q x_k w)}{(1 - x_k w)}\; .
\end{equation}
Note that from this expression, we cannot decompose this function as
in~(\ref{eq:h-exp}) since the coefficients in this expansion depend on
\(\bm{x}\) and \(\bm{y}\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Big Schur functions expansion}
Now, let's revisit the result originally derived in~\cite{Foda:2008hn}
which demonstrates that the scalar product~(\ref{eq:scalar-hl}) is a
tau function of the KP hierarchy. According to~\cite[Chapter 3,
  Section 4, Equation (4.7)]{Macdonald:1998}, we can expand the Cauchy
identity~(\ref{eq:cauchy_hl}) as
\begin{equation}
 \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k} = 
\sum_{\lambda} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
\sum_{\lambda} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; .
\end{equation}
Here, the polynomials \(\mathcal{S}_\lambda\), known as the big-Schur
functions, are defined by a Jacobi-Trudi formula:
\begin{equation}
  \mathcal{S}_{\lambda} (\bm{y}, Q) = \det(q_{\lambda_i -i + j}(\bm{y}, Q))\; , 
\end{equation}
and the coefficients \(q_m\) are obtained from the expression:
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_{j} \frac{1-Q y_j z}{1 - y_j z}\; .
\end{equation}
where \(z\) is a formal variable. It has been
argued~\cite{Foda:2008hn} in that if we interpret the big-Schur
functions as Plücker coordinates, the expression
\begin{equation}
\label{eq:big-expansion}
  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})
  = \sum_{\lambda \subseteq [N,M]} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y})
  = \sum_{\lambda \subseteq [N,M]} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x})\; ,
\end{equation}
is a restricted KP tau function with respect to both set of coordinates,
that is \(\bm{x}\) and \(\bm{y}\).

Based on these findings, we conclude that the inner product can be
expressed as a quotient of two Toda tau functions or as a KP tau
function with coefficients given by the big Schur
polynomials. However, further exploration of these results is
warranted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Kostka-Foulkes expansion}
In a Schur polynomial basis, we write
\begin{equation}
P_\lambda(\bm{x}, Q) = \sum_{\mu} K^{-1}_{\lambda \mu}(Q) s_\lambda(\bm{x})\; , 
\end{equation}
where \(K^{-1}_{\mu\nu}(Q) \in \mathbb{Z}[Q]\) are inverse
Kostka-Foulkes polynomials~\cite{Macdonald:1998, Wheeler:2018}.

If we now insert this expansion in the inner product, we have 
\begin{equation}
    \sum_\lambda b_\lambda(Q) P_\lambda(\bm{x}, Q) P_\lambda(\bm{y}, Q) = 
    \sum_{\mu,\nu} \left(\sum_\lambda K_{\mu \lambda}^{-1 \; T} b_\lambda(Q)  K_{\lambda \nu}^{-1} \right)
    s_{\mu}(\bm{x}) s_{\nu}(\bm{y}) 
\end{equation}
and we have the coefficients 
\begin{equation}
\label{eq:c-expan}
 \tilde{c}_{\mu\nu}(Q) = \sum_\lambda K_{\mu \lambda}^{-1 \; T} b_\lambda(Q)  K_{\lambda \nu}^{-1} \; ,
\end{equation}
for the double expansion of the inner product in terms of Schur
polynomials. Based on our discussion so far, we know that these
coefficients are not Plücker coordinates of the Toda hierarchy.

But we can say something interesting about these coefficients. Let us
now expand the big-Schur functions as
\begin{equation}
\label{eq:kf-exp}
  \mathcal{S}_\mu(\bm{x}, Q) = \sum_{\lambda} \mathcal{C}^{\mu}_\lambda(Q) s_\lambda(\bm{x})\; ,
\end{equation}
we can fix the coefficients \(\mathcal{C}_\lambda^\mu(Q)\) in terms of
Kostka-Foulkes polynomials using the orthogonality relations of these
polynomials~\cite{Macdonald:1998}. In particular, there is an inner
product in the ring of symmetric functions such that
\begin{equation}
  b_\mu \langle P_\mu(\bm{x}, Q), P_\nu(\bm{x}, Q)\rangle = \langle
  \mathcal{S}_\mu(\bm{x}, Q), s_\nu(\bm{x})\rangle = \delta_{\mu\nu}\; .
\end{equation}
Therefore,
\begin{equation}
  \begin{split}
\delta_{\mu\nu} & = \sum_\lambda \mathcal{C}_\lambda^\mu \langle s_\lambda(\bm{x}), s_\nu(\bm{x})\rangle
 = \sum_\lambda \sum_{\pi, \sigma} \mathcal{C}_\lambda^\mu K_{\lambda \pi} K_{\nu \sigma}
\langle P_{\pi}(\bm{x}, Q), P_{\sigma}(\bm{x}, Q)\rangle \\ 
& = \sum_\lambda \mathcal{C}_\lambda^\mu \sum_{\sigma}K_{\lambda \sigma} b_{\sigma}^{-1} K_{\nu \sigma}\; ,
  \end{split}
\end{equation}
and we conclude that 
\begin{equation}
\label{eq:indices}
(\mathcal{C}^{-1})^\lambda_\nu = \sum_\sigma K_{\lambda\sigma} b^{-1}_\sigma (K_{\sigma\nu})^T\; ,
\end{equation}
and consistency with~(\ref{eq:c-expan}) implies that
\(\mathcal{C}^\lambda_\nu \equiv \tilde{c}_{\lambda\nu}\).

Hence, we conclude that 
\begin{equation}
S_{\mu}(\bm{t},Q) = \sum_{\nu} \tilde{c}_{\mu\nu}(Q) s_{\mu}(\bm{t})\; ,
\end{equation}
where we have written the big Schur polynomial in terms of the Miwa
coordinates \(\bm{t}\).

\begin{remark}
This expression also reveals something interesting. We can formulate
this problem in terms of partitions defined in the phase
model. Alternatively, we can utilize the conventional vertex operator
construction rather than the q-deformed version proposed by
Jing~\cite{Jing1991, Jing1995}, which is much more challenging to
handle.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Supersymmetric Schur polynomials expansion}
With this result, we now argue that although~(\ref{eq:scalar-hl}) is
not a Toda tau function with respect to the coordinates \(\{\bm{t};
\bm{t}'\}\), we can define a new set of coordinates
\(\{\bm{t};\bm{T}\}\) that render the inner product a trivial Toda tau
function.

Let us first decompose
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_j (1 + y_j (-Q z)) \prod_k (1 - y_k z)^{-1} = 
 \sum_{j, k} e_j(\bm{y}) h_k(\bm{y}) (- Q)^j z^{j+k}\; .
\end{equation}
Reorganizing this sum, we conclude that
\begin{equation}
  q_m(\bm{y}, Q)  = \sum_{j=0}^m e_j(\bm{y}) h_{m-j}(\bm{y}) (- Q)^j =
 \sum_{j=0}^m e_j(-Q\bm{y}) h_{m-j}(\bm{y}) \; ,
\end{equation}
where in the last equality we have used the homogeneity of the
elementary symmetric polynomials, and \( Q\bm{y} \equiv (Qy_1, Qy_2,
\dots)\).

It is easy to see that
\begin{subequations}
\begin{equation}
  \begin{split}
    \prod_{j=1}^N (1 - (Q y_j)z) & = \prod_{j=1}^N e^{ \ln  (1 - (Q y_j)z)} = 
    \prod_{j=1}^N \exp \left( - \sum_{n>0} \frac{Q^n y^n_j}{n} z^n \right) \\ 
    & = \exp \left( - \sum_{n>0} \sum_{j=1}^N \frac{Q^n y^n_j}{n} z^n \right)  =
    \exp \left( - \sum_{n>0} t^{(Q)}_n z^n \right)  
  \end{split}
\end{equation}
where \(t^{(Q)}_n = \frac{Q^n}{n} \sum_j y_j^n = Q^n t'_n\). Additionally
\begin{equation}
    \prod_{j=1}^N \frac{1}{(1 - y_jz)} = \prod_{j=1}^N e^{- \ln  (1 - y_jz)} = 
    \prod_{j=1}^N \exp \left( \sum_{n>0} \frac{y_j}{n} z^n \right)=
    \exp \left(\sum_{n>0} t_n z^n \right) \; . 
\end{equation}
\end{subequations}

Consequently
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_j \frac{1 - Q y_j z}{1 - y_k z}
 = \exp \left( \sum_{n\geq 1} (t'_n - t_n^{(Q)})z^n \right)
 \equiv \exp \left( \sum_{n\geq 1} T_n z^n \right)\; ,
\end{equation}
with \(T_n = (1 - Q^n) t'_n\). Then, we conclude that \(q_m(\bm{y}, Q)\)
are homogeneous polynomials with respect the Miwa coordinates
\(\bm{T} = (T_1, T_2, \dots)\). Then
\begin{equation}
  \mathcal{S}_\lambda(\bm{t}', Q) = \det \left(h_{\lambda_i - i +j}(\bm{T})\right) = s_\lambda(\bm{T})\; .
\end{equation}
All in all, we conclude that the big Schur functions
\(S_\lambda(\bm{t}', Q)\) are ordinary Schur functions with respect to
the coordinates \(\bm{T}\).

Moreover, a more refined approach is also possible. Supersymmetric (or
Hook) Schur functions~\cite{Berele:1983}, as discussed in works such
as~\cite{Macdonald:1998, Moens:2003}, denoted by
\(s_\lambda(\bm{\alpha}/\bm{\beta})\), are defined as ordinary Schur
functions evaluated at Miwa coordinates of the form
\begin{equation}
  T_n = \frac{1}{n} \left( \sum_{i=1}^{\dim(\alpha)} \alpha_i^n
       - \sum_{i=1}^{\dim(\beta)} (-\beta_i)^n\right)\; .
\end{equation}
Comparing this expression with the results above, we can see that the
big Schur functions \(S_\lambda(\bm{t}', Q)\) correspond to the
supersymmetric Schur functions for \(\bm{\alpha} = \bm{y}\) and
\(\bm{\beta} = - Q \bm{y}\), which is
\begin{equation}
  S_\lambda(\bm{t}', Q) = s_\lambda[\bm{y}/(- Q\bm{y})]\; .
\end{equation}

Putting these facts together, we immediately conclude that
\begin{equation}
\lim_{M,N\to \infty}\mathcal{I}_Q(N,M|\bm{x}, \bm{y}) = 
  \sum_\lambda s_\lambda(\bm{T}) s_\lambda (\bm{t})
\end{equation}
is a Toda hierarchy tau function with respect to
\(\{\bm{t};\bm{T}\}\). As a result, we obviously have that
\begin{equation}
  \mathcal{I}_Q(N,M| \bm{x}, \bm{y}) \equiv \mathcal{I}_Q(N,M| \bm{t}, \bm{T})
  = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{T}) s_\lambda(\bm{t})
\end{equation}
is also a restricted tau function of the Toda hierarchy with respect
to \(\bm{t}\) and \(\bm{T}\).

Let us also express the supersymmetric Schur polynomials in terms of
ordinary Schur functions, as shown in~\cite[Sec. I.5, exerc. 23]{Macdonald:1998}:
\begin{equation}
  s_{\lambda}(\bm{x}/\bm{y}) = \sum_\mu s_{(\lambda/\mu)'}(\bm{y}) s_\mu(\bm{x})\; ,
\end{equation}
where the prime denotes the conjugate diagram. Compare this expansion
with~(\ref{eq:kf-exp}). Then
\begin{equation}
  \label{eq:tau-hl}
  \sum_{\lambda } s_\lambda(\bm{T}) s_\lambda(\bm{t}) =
  \sum_{\lambda,\mu } s_{(\lambda/\mu)'}(- Q\bm{y}) s_\lambda(\bm{t}) s_\mu(\bm{t}')\; .
\end{equation}

From this expression we conclude (and speculate) the following.

\begin{remark}
Since the skew Schur polynomials have determinant expressions, we can
deduce that the \(\bm{y}\)-dependent coefficients \(c_{\mu \lambda} =
s_{(\lambda/\mu)'}(-Q\bm{y})\) have Jacobi-Trudi expressions. It is
tempting to regard these objects as \(\bm{y}\)-dependent Plücker
coordinates. In this sense, we would have a curve in the infinite
Grassmannian instead of a point.
\end{remark}

\begin{remark}
It is worth noting that one of the simplest nontrivial solutions of
the KP hierarchies are the Schur functions
themselves~\emph{\cite{Zabrodin2018}}. As we have just demonstrated, the big
Schur polynomials can be expressed as supersymmetric Schur
polynomials, which are essentially ordinary Schur polynomials in a
specific choice of Miwa coordinates. Therefore, we conclude that the
big Schur polynomials are also KP tau functions. This direct
conclusion serves as an alternative proof of this fact, originally derived in
in~\emph{\cite{Necoechea:2019wbg}} using the KP bilinear identity.
\end{remark}

Combining these observations, we conclude that the left-hand side of
equation (\ref{eq:tau-hl}) serves as a tau function with respect to
the coordinates \(\{\bm{t}; \bm{T}\}\). Additionally, utilizing the
vertex operator formalism, we can express the Schur polynomials as
\begin{equation}
  s_{\lambda}(\bm{t}') = \bra{\lambda} e^{J_-(\bm{t}')}\ket{\bm{0}}\qquad
  s_{\mu}(\bm{t}) = \bra{\bm{0}} e^{J_+(\bm{t})}\ket{\mu}\; ,
\end{equation}
we have
\begin{equation}
  \sum_{\lambda,\mu } c_{\mu\lambda}(Q,\bm{y}) s_\mu(\bm{t}') s_\lambda(\bm{t}) = 
  \sum_{\lambda,\mu } \bra{\bm{0}} e^{J_+(\bm{t})}\ket{\mu} c_{\mu\lambda}(Q,\bm{y})
  \bra{\lambda} e^{J_-(\bm{t}')}\ket{\bm{0}} \; .
\end{equation}
It is important to emphasize once again that this expression does not
constitute a tau function with respect to \(\{\bm{t},
\bm{t}'\}\). Despite this crucial distinction, we proceed under the
assumption that the coefficients can be expressed as
\begin{equation}
  c_{\mu\lambda}(Q, \bm{y}) = \bra{\mu} G_Q(\bm{y}) \ket{\lambda}\; ,
\end{equation}
where \(G_Q(\bm{y}) \in GL(\infty)\). Hence
\begin{equation}
  \lim_{N,M\to\infty} \mathcal{I}_Q(N, M| \bm{t}, \bm{t}'(\bm{y})) = 
  \bra{\bm{0}} e^{J_+(\bm{t})} G_Q(\bm{y}) e^{J_-(\bm{t}')}\ket{\bm{0}}\; .
\end{equation}
where \(t'_n = \frac{1}{n}\sum_i y_i^n\).

To derive the expression for this group element, recall that \(T_m =
t'_m - Q^m t'_m\). With this in mind, we can express the scalar
product as follows:
\begin{subequations}
\begin{equation}
  \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
\bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{t}')  - J_-(\bm{t}^{(Q)})} \ket{\bm{0}} \; .
\end{equation}
From the Heisenberg algebra, we deduce that \([J_-(\bm{t}),
  J_-(\bm{t}')] = 0\), therefore
\begin{equation}
  \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
\bra{\bm{0}} e^{J_+(\bm{t})} e^{-J_-(\bm{t}^{(Q)})} e^{J_-(\bm{t}')}  \ket{\bm{0}} \; .
\end{equation}
\end{subequations}
We finally conclude that 
\begin{equation}
  G_{Q}(\bm{y}) = \exp \left(-J_-(\bm{t}^{(Q)})\right) \qquad
  t^{(Q)}_n = \frac{Q^n}{n}\sum_j y_j^n\; . 
\end{equation}

Generally, the expectation values involving coordinate-dependent group
elements \(G_{Q}(\bm{y})\) do not form Toda hierarchy tau
functions. However, in the example above, the coordinates combine in
such a way that they generate a tau function with respect to the
coordinates \(\bm{T}\) and \(\bm{t}\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Conclusions and Perspectives}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

In this work, we have explored the connections between a quantum
integrable system, the q-boson model, and solutions of a classical
integrable system, the Toda hierarchy. Our investigation has extended
some early findings in this field and has also unveiled new research
avenues, which we aim to explore in future studies.  Let us now delve
into some of these promising directions.

In Section 3, we delved into various facets of the phase model. We
presented the scalar product of two off-shell Bethe states as an
elementary example of the Toda system tau function. Additionally, we
revealed an intriguing alternative portrayal of these scalar products
using a matrix integral framework, which corresponds to an ensemble of
Coulomb particles. This discovery opens up promising avenues for
further inquiry. Specifically, it would be interesting to delve deeper
into this subject and explore whether the matrix model description
offers valuable insights into the phase model and its Bethe
roots. Such an investigation promises to illuminate the underlying
dynamics of the phase model and its relationship with classical
integrable systems.

We have elucidated how various correlation functions align with this
framework, revealing a remarkably rich structure. Notably, we explored
a mapping between the phase model data and the vertex operator
representation of free bosons. An intriguing avenue for future
research is to delve into the properties of the hypergeometric tau
functions uncovered in Section 3.3.1. Investigating these functions
promises to yield valuable insights, and comparing them with existing
literature will further enhance our understanding of their
characteristics and implications within the context of integrable
systems.

In Section 4, we addressed the same problem within the context of the
q-boson model. Initially, we derived a determinant formula for the
q-boson, followed by a discussion on its expression as a quotient of
two phase model tau functions. This result establishes a connection
between the q-boson and phase model scalar product. Furthermore, we
explored three different expansions for these scalar products. Notably, we
observed that they can be expanded in terms of Big Schur and
supersymmetric tau functions. Consequently, we introduced a new set of
coordinates and demonstrated that it can be expressed as a Toda tau
function in terms of this new coordinate set.

One of our most pressing challenges lies in elucidating the intricate
connection between our findings and the Ablowitz-Ladik
hierarchy. Since this hierarchy arises as a reduction of the Toda
hierarchy, it becomes imperative to understand how we can capture the
structure of the former using the results we've derived. Given that
the q-boson model can be seen as a quantization of the Ablowitz-Ladik
equation, it stands to reason that we should observe some resemblance of
this classical problem within the framework of the q-boson
system. Understanding this relationship promises valuable insights
into the interplay between classical and quantum integrable systems.

We hope to address some of these challenges in future publications.

\subsubsection*{Acknowledgments}
This work is supported by Fapesp through grant \textbf{2022/06599-0}.

\printbibliography

\end{document}