### 2024-06-22 referee-answer.txt
answer to referee

Dear Editor,

Below are my responses to the referee's assessment:

Dear Referee,

I agree that my results are somewhat buried in the manuscript. Let me
start by saying that I was very careful not to claim credit for any
previous derivations. Some discussions in my current paper have been
addressed before, but not clearly enough. Therefore, this general
exposition also highlights my work and the field in
general. Additionally, there is pedagogical value in this.

Now, let me explain my main new results.

Section 3:

This section has more overlap with existing literature, particularly
with a paper by Wheeler that I cite in my work. In his paper, Wheeler
shows that the inner product of two Bethe states are KP hierarchy tau
functions. He finds the coefficients of the expansion and demonstrates
that these coefficients satisfy Plücker relations. My Section 3.1
improves on this result. I argue that besides the KP hierarchy, these
scalar products are also Toda hierarchy tau functions, as they can be
expanded as a Cauchy identity, with the coefficients being Schur
functions (see equation (28)). If we fix one set of coefficients in
this expression, Wheeler's results are recovered, as argued in
equation (40).

Section 3.2 is new and considers correlation functions. I wanted to
determine when the correlation functions satisfy integrable hierarchy
equations. Using expressions derived by Bogoliubov (see eq. (42.a)), I
rewrote them in the appropriate form to check if the correlation
functions are KP/Toda tau functions. For example, (42.b) cannot be
written as a Toda tau function, but with some adjustments, it can be
written as (51), which is a KP tau function. From the expansion (52),
also derived by Bogoliubov, we find (56) and (57), showing that these
expressions generally do not satisfy integrable hierarchy equations.

Section 3.3 presents my main results. Here, I argue that the general
form of tau functions associated with this phase model must be (67),
where G is given by (63). I also mention that from this general
expression, we can consider interesting cases, such as (68), which
fall into the class of hypergeometric tau functions. These problems
are currently under investigation in my follow-up paper.

Section 4:

This section attempts to replicate the analysis done in Section
3. There aren't many known results in the literature regarding the
relation between these objects and KP/Toda tau functions. I use known
formulas to see if they are tau functions. Wheeler has shown that
scalar products can be viewed as KP tau functions, as before.

I then write these formulas in terms of simpler orthogonal
polynomials. Firstly, I show that these scalar products cannot be
naively written as Toda tau functions (see (84)). However, using some
known expansions (see (85)), we can expand the inner products in
different forms. In one of them, we define a new set of coordinates
(equation (102)), making the scalar product of Q-Bosons Toda tau
functions with respect to these new coordinates. This is my main
result in this section. Additionally, I propose a conjecture worth
investigating in the future, as mentioned in Remark 2, regarding the
existence of 'functional' Plücker coordinates in the infinite
Grassmannian.

My Remark 3 also presents a new result. I emphasize that I do not wish
to take undue credit. Reference [NR19] uses an intricate construction
to show that the Big Schur functions are KP tau functions. Although
significant, my results trivialize their analysis. Since a simple
coordinate transformation makes the Big Schur functions ordinary Schur
functions (see equation (101)), this fact is trivial. I have included
this as a remark rather than stating it bluntly.

Having said that, let me return to my previous assessment of other
references. I am very interested in this field but have noticed that
papers on these topics are scattered. For example, in section 3.1.1,
the relation between the phase model and the ensemble of 2D Coulomb
particles seems obvious, but there are no references. I hope this
paper will draw attention to this particular problem. As I mentioned,
I am currently exploring this topic, but it is more challenging than I
previously thought.

Overall, I do not view this work as a full collection of new results
but as setting the stage for a research field that is somewhat
inactive. I use previous results to organize and draw attention to
open problems. In this process, I prove some new results, which,
though not major, serve as a starting point to investigate those open
problems.

I have added comments to my conclusions and improved my introduction.

I hope these clarifications are helpful.

Best regards,
Thiago